Building the Million Transcriptome Ocular Meta-Atlas

### 2021-01-17 manuscript_00.Rmd
updates

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_meta']] <- 0
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['label_toggle']] <- 2
#input[['meta_filter_cat']] <- 'CellType_predict'
#input[['meta_filter_on']] <- 'Bipolar Cells'
ctp <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_meta']] <- 1
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- c('AC/HC Precursors','Amacrine Cells','Astrocytes','Bipolar Cells','Cones','Early RPCs','Endothelial','Horizontal Cells','Late RPCs','Microglia','Muller Glia','Neurogenic Cells','Pericytes','PR Precursors','Red Blood Cells','Retinal Ganglion Cells','Rod Bipolar Cells','Rods','RPCs','RPE','Schwann','Tabula Muris')
ctp2 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_meta']] <- 0
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
#input[['meta_filter_cat']] <- 'CellType_predict'
#input[['meta_filter_on']] <- 'Bipolar Cells'
org <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

@@

First, we project another retina dataset onto scEiaD from Fadl et al . As this is a pan-retina tissue sample, we would expect most of the major cell types of the retina to be present. In the Google colab notebook we show how a set of the reads from SRR12130660 can be processed from the raw reads to a UMAP visualization overlapping our scEiaD in under 15 minutes. Indeed, as expected we see many amacrine, rods, muller glia, bipolar, a few cones and retinal ganglion, and a small number of non-retina cells (`r supFig_cap(name = 'projection_fig', display = 'cite')`b).

@@

In order to identify cell types for the `r meta_filter %>% filter(is.na(CellType), is.na(TabulaMurisCellType))` unlabelledcells we designed a custom xgboost based cell type classifier. We took labeled data and split it into training and test sets, stratified by cell type. The input features used to train the model are the scVI latent dimensions, the total number of reads in each cell, the number genes detected in each cell, and the percent mitochondrial gene expression of each cell. We additionally generated features using the age of each sample by group sample into three developmental categories (Early Development, Late Development, and Adult) and then generated a one-hot encoded feature for each category. In order to speed up training times, we used the gpu implementation of the xgboost algorithm from the the xgboost python library. The model was trained using default parameters. The trained model had an overall F1 score of XXX an average precision and recall of XXX and XXX for each major retinal cell type. This model was then used to identify labels for each unlabelled cell. Unlabelled data was preprocessed identically to training data and fed into model to generate a vector of label probabilities for each cell. We selected the highest label probability for each cell, and required a minimum probability of .5 to assign a label to a cell.

@@

```{r supFig_platformInt, fig.width=12, fig.height=10,  fig.cap=supFig_platformInt_cap, echo=FALSE, message = FALSE}
supFig_platformInt_cap <- supFig_cap(name='supFig_platformInt', caption =  '')
input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_meta']] <- 5
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
#input[['meta_filter_cat']] <- 'CellType_predict'
#input[['meta_filter_on']] <- 'Bipolar Cells'
ctp3 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

@@

>>>>

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 0
input[['pt_size_meta']] <- 0
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['label_toggle']] <- 2
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- NULL
ctp <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 1
input[['pt_size_meta']] <- 1
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- c('AC/HC Precursors','Amacrine Cells','Astrocytes','Bipolar Cells','Cones','Early RPCs','Endothelial','Horizontal Cells','Late RPCs','Microglia','Muller Glia','Neurogenic Cells','Pericytes','PR Precursors','Red Blood Cells','Retinal Ganglion Cells','Rod Bipolar Cells','Rods','RPCs','RPE','Schwann','Tabula Muris')
ctp2 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 0
input[['pt_size_meta']] <- 0
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- NULL
org <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

@@

First, we project another retina dataset onto scEiaD from Fadl et al. As this is a pan-retina tissue sample, we would expect most of the major cell types of the retina to be present. In the Google colab notebook we show how a set of the reads from SRR12130660 can be processed from the raw reads to a UMAP visualization overlapping our scEiaD in under 15 minutes. Indeed, as expected we see many amacrine, rods, muller glia, bipolar, a few cones and retinal ganglion, and a small number of non-retina cells (`r supFig_cap(name = 'projection_fig', display = 'cite')`b).

@@

In order to identify cell types for the `r format(meta_filter %>% filter(is.na(CellType), is.na(TabulaMurisCellType)) %>% nrow(), big.mark=",", scientific=FALSE)` unlabelledcells we designed a custom xgboost based cell type classifier. We took labeled data and split it into training and test sets, stratified by cell type. The input features used to train the model are the scVI latent dimensions, the total number of reads in each cell, the number genes detected in each cell, and the percent mitochondrial gene expression of each cell. We additionally generated features using the age of each sample by group sample into three developmental categories (Early Development, Late Development, and Adult) and then generated a one-hot encoded feature for each category. In order to speed up training times, we used the gpu implementation of the xgboost algorithm from the the xgboost python library. The model was trained using default parameters. The trained model had an overall F1 score of XXX an average precision and recall of XXX and XXX for each major retinal cell type. This model was then used to identify labels for each unlabelled cell. Unlabelled data was preprocessed identically to training data and fed into model to generate a vector of label probabilities for each cell. We selected the highest label probability for each cell, and required a minimum probability of .5 to assign a label to a cell.

@@

```{r supFig_platformInt, fig.width=12, fig.height=10,  fig.cap=supFig_platformInt_cap, echo=FALSE, message = FALSE}
supFig_platformInt_cap <- supFig_cap(name='supFig_platformInt', caption =  '')
input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 5
input[['pt_size_meta']] <- 5
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
#input[['meta_filter_cat']] <- 'CellType_predict'
#input[['meta_filter_on']] <- 'Bipolar Cells'
ctp3 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

@@

```{r projection_fig, fig.width=12, fig.height=10,  fig.cap=projection_fig_cap, echo=FALSE, message = FALSE}
projection_fig_cap <- supFig_cap(name='projection_fig', caption =  'bloop')
1 + 1
```
